% =========================
% chapters/04_phaser_vs_urukul.tex
% =========================
\chapter{Choosing the RF Engine: Phaser vs Urukul, and DDS vs AWG}
\label{ch:phaser_vs_urukul}

\section{Scope of comparison}
Both Urukul and Phaser integrate into ARTIQ/Sinara crates, but they represent distinct signal-generation
philosophies: DDS-centric synthesizer (Urukul) versus AWG-class DAC with interpolation, mixing, and
gateware-defined waveform modes (Phaser). This chapter compares them along dimensions that matter for
trapped-ion AOM and gate control: phase control, update timing, bandwidth, synchronization pathways, and
waveform flexibility.

\section{Specification-driven comparison}
\input{tables/phaser_urukul_comparison}

\section{Interpretation: when Phaser is “better” and when Urukul is “better”}
Urukul excels when the experiment primarily needs agile sinusoidal tones with extremely fine frequency
resolution and well-characterized, phase-coherent updates (especially with careful IO\_UPDATE alignment and
SYNC tree configuration, when used). Phaser excels when the experiment needs (i) wideband or shaped
waveforms, (ii) simultaneous multitone generation, or (iii) multi-stage digital mixing/interpolation workflows
that more closely resemble a modern transmitter chain.

\section{DDS vs AWG: a control-systems viewpoint}
In a DDS, a phase accumulator and lookup/cordic stage generate a sinusoid referenced to a high-quality
clock, offering fine frequency and phase programmability with predictable spurious behavior.
In an AWG, user-defined samples are played from memory at a set sample rate; AWGs are therefore highly
flexible but require careful attention to sample clocks, reconstruction filters, and memory/sequence
constraints. Many commercial instruments explicitly distinguish these paradigms and use cases.

For trapped-ion control, the practical distinction is this:
\emph{DDS is ideal for a few clean tones with tight frequency control; AWG-class generation is ideal for
pulse shaping, modulation, and many-tone composite waveforms.}
